---
title: "Replication material: Gneiting and Vogel (2021)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

The following scripts replicate the results shown in Table 1 and Figure 1 from the study by Gneiting and Vogel (2021).

```{r message=FALSE, warning=FALSE}
library(tidyverse)

devtools::load_all()

data("aSAH_Robin")
data("hiv_Sing")
data("PSA_Etzioni")
data("WS_Vogel")

sessionInfo()

```

```{r}
empROC <- lapply(
  list(aSAH_Robin, hiv_Sing, PSA_Etzioni, WS_Vogel),
  function(x) {
    roc(obs ~ forc, x)
  })

MDE_info <- list(
  list(method = "bin2p", info = "unrestricted"),
  list(method = "bin2p", info = "concave"),
  list(method = "beta2p", info = "unrestricted"),
  list(method = "beta2p", info = "concave")
)

df <- tibble(
  study = rep(c("Robin", "Sing", "Etzioni", "Vogel"), each = 4),
  empROC = rep(empROC, each = 4),
  MDE_info = rep(MDE_info, times = 4)
) %>%
  mutate(
    params = map2(empROC, MDE_info, MDE)
  )

```

The following script file runs the Monte Carlo goodness-of-fit test for each data set and parametric model, and saves the results in the subfolder "goodness_of_fit_files". For the smaller data sets the runtime can be less than one minute, whereas the runtime for the larger data sets can be up to around one hour. These runtimes have been recorded on notebook with an Intel(R) Core(TM) i7-10810U processor.

```{r eval = FALSE}
source("goodness_of_fit.R")

```


```{r}
df$pvalue <- map2(df$study, df$MDE_info, function(study, MDE_info) {
  filename <- sprintf("goodness_of_fit_files/%s_%s_%s.RData",
                      study, MDE_info$method, MDE_info$info)
  load(filename)
  results$pval
})

```

The following functions are used to print and plot the results.

```{r}
format_results <- function(df) {
  with(df, {
    c(
      sprintf("Binormal model - unrestricted"),
      sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
              params[[1]]$pars_fit[1], params[[1]]$pars_fit[2]),
      sprintf("L2-distance:                 %.3f",
              params[[1]]$L2_fit),
      sprintf("P-value:                     %.3f",
              pvalue[1]),
      
      sprintf("Binormal model - concave"),
      sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
              params[[2]]$pars_fit[1], params[[2]]$pars_fit[2]),
      sprintf("L2-distance:                 %.3f",
              params[[2]]$L2_fit),
      sprintf("P-value:                     %.3f",
              pvalue[2]),
      
      sprintf("Beta model - unrestricted"),
      sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
              params[[3]]$pars_fit[1], params[[3]]$pars_fit[2]),
      sprintf("L2-distance:                 %.3f",
              params[[3]]$L2_fit),
      sprintf("P-value:                     %.3f",
              pvalue[3]),
    
      sprintf("Beta model - concave"),
      sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
              params[[4]]$pars_fit[1], params[[4]]$pars_fit[2]),
      sprintf("L2-distance:                 %.3f",
              params[[4]]$L2_fit),
      sprintf("P-value:                     %.3f",
              pvalue[4])
    )
  }) %>%
    cat(sep = "\n")
}

plot_results <- function(df_study) {
  with(df_study, {
    p <- plot_roc_empirical(empROC[[1]])
    p <- plot_binormal(params[[1]]$pars_fit, MDE_info[[1]], p,
                       color = "red")
    p <- plot_binormal(params[[2]]$pars_fit, MDE_info[[2]], p,
                       lty = 2, color = "red")
    p <- plot_beta(params[[3]]$pars_fit, MDE_info[[3]], p,
                   color = "blue")
    p <- plot_beta(params[[4]]$pars_fit, MDE_info[[4]], p,
                   lty = 2, color = "blue")
    p +
      theme(aspect.ratio = 1) +
      geom_segment(mapping = aes(x = 0, xend = 1, y = 0, yend = 1), lty = 2)
  })
}
```



### Etzioni et al. (1999)

Prostate cancer antigen ratio, 116 observations

```{r EtzioniEtAl1999, warning = FALSE}
df_Etzioni <- filter(df, study == "Etzioni")

format_results(df_Etzioni)
plot_results(df_Etzioni)

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.


### Sing et al. (2005)

Coreceptor usage SVM predictor, 3450 observations

```{r SingEtAl2005}
df_Sing <- filter(df, study == "Sing")

format_results(df_Sing)
plot_results(df_Sing)

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.

### Robin et al. (2011)

Clinical outcome S100$\beta$ concentration, 113 observations

```{r RobinEtAl2011}
df_Robin <- filter(df, study == "Robin")

format_results(df_Robin)
plot_results(df_Robin)


```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.


### Vogel et al. (2018)

Precipitation NWP forecast, 5449 observations

```{r VogelEtAl2018}
df_Vogel <- filter(df, study == "Vogel")

format_results(df_Vogel)
plot_results(df_Vogel)

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.
