---
title: "Replication material: Gneiting and Vogel (2021)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(devtools)
load_all()

data("aSAH_Robin")
data("hiv_Sing")
data("PSA_Etzioni")
data("WS_Vogel")
```

```{r}
empROC <- lapply(
  list(aSAH_Robin, hiv_Sing, PSA_Etzioni, WS_Vogel),
  function(x) {
    roc(obs ~ forc, x)
  })

MDE_info <- list(
  list(method = "bin2p", info = "unrestricted"),
  list(method = "bin2p", info = "concave"),
  list(method = "beta2p", info = "unrestricted"),
  list(method = "beta2p", info = "concave")
)

df <- tibble(
  study = rep(c("Robin", "Sing", "Etzioni", "Vogel"), each = 4),
  empROC = rep(empROC, each = 4),
  MDE_info = rep(MDE_info, times = 4)
) %>%
  mutate(
    params = map2(empROC, MDE_info, fit_MDE)
  )

```

### Etzioni et al. (1999)

Prostate cancer antigen ratio, 116 observations

```{r warning = FALSE}
df_Etzioni <- filter(df, study == "Etzioni")

with(df_Etzioni, {
  c(
    sprintf("Binormal model - unrestricted"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[1]]$pars_fit[1],
            params[[1]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[1]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[2]]$pars_fit[1],
            params[[2]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[2]]$L2_fit),
    
    sprintf("Beta model - unrestricted"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[3]]$pars_fit[1],
            params[[3]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[3]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[4]]$pars_fit[1],
            params[[4]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[4]]$L2_fit)
  )
}) %>%
  cat(sep = "\n")

with(df_Etzioni, {
  p <- plot_roc_empirical(empROC[[1]])
  p <- plot_binormal(params[[1]]$pars_fit, MDE_info[[1]], p, color = "red")
  p <- plot_binormal(params[[2]]$pars_fit, MDE_info[[2]], p, lty = 2, color = "red")
  p <- plot_beta(params[[3]]$pars_fit, MDE_info[[3]], p, color = "blue")
  p <- plot_beta(params[[4]]$pars_fit, MDE_info[[4]], p, lty = 2, color = "blue")
  p +
    theme(aspect.ratio = 1) +
    geom_segment(mapping = aes(x = 0, xend = 1, y = 0, yend = 1), lty = 2)
})

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.


### Sing et al. (2005)

Coreceptor usage SVM predictor, 3450 observations

```{r echo = FALSE}
df_Sing <- filter(df, study == "Sing")

with(df_Sing, {
  c(
    sprintf("Binormal model - unrestricted"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[1]]$pars_fit[1],
            params[[1]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[1]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[2]]$pars_fit[1],
            params[[2]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[2]]$L2_fit),
    
    sprintf("Beta model - unrestricted"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[3]]$pars_fit[1],
            params[[3]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[3]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[4]]$pars_fit[1],
            params[[4]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[4]]$L2_fit)
  )
}) %>%
  cat(sep = "\n")

with(df_Sing, {
  p <- plot_roc_empirical(empROC[[1]])
  p <- plot_binormal(params[[1]]$pars_fit, MDE_info[[1]], p, color = "red")
  p <- plot_binormal(params[[2]]$pars_fit, MDE_info[[2]], p, lty = 2, color = "red")
  p <- plot_beta(params[[3]]$pars_fit, MDE_info[[3]], p, color = "blue")
  p <- plot_beta(params[[4]]$pars_fit, MDE_info[[4]], p, lty = 2, color = "blue")
  p +
    theme(aspect.ratio = 1) +
    geom_segment(mapping = aes(x = 0, xend = 1, y = 0, yend = 1), lty = 2)
})

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.

### Robin et al. (2011)

Clinical outcome S100$\beta$ concentration, 113 observations

```{r echo = FALSE}
df_Robin <- filter(df, study == "Robin")

with(df_Robin, {
  c(
    sprintf("Binormal model - unrestricted"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[1]]$pars_fit[1],
            params[[1]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[1]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[2]]$pars_fit[1],
            params[[2]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[2]]$L2_fit),
    
    sprintf("Beta model - unrestricted"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[3]]$pars_fit[1],
            params[[3]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[3]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[4]]$pars_fit[1],
            params[[4]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[4]]$L2_fit)
  )
}) %>%
  cat(sep = "\n")

with(df_Robin, {
  p <- plot_roc_empirical(empROC[[1]])
  p <- plot_binormal(params[[1]]$pars_fit, MDE_info[[1]], p, color = "red")
  p <- plot_binormal(params[[2]]$pars_fit, MDE_info[[2]], p, lty = 2, color = "red")
  p <- plot_beta(params[[3]]$pars_fit, MDE_info[[3]], p, color = "blue")
  p <- plot_beta(params[[4]]$pars_fit, MDE_info[[4]], p, lty = 2, color = "blue")
  p +
    theme(aspect.ratio = 1) +
    geom_segment(mapping = aes(x = 0, xend = 1, y = 0, yend = 1), lty = 2)
})

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.


### Vogel et al. (2018)

Precipitation NWP forecast, 5449 observations

```{r echo = FALSE}
df_Vogel <- filter(df, study == "Vogel")

with(df_Vogel, {
  c(
    sprintf("Binormal model - unrestricted"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[1]]$pars_fit[1],
            params[[1]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[1]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (mu, sigma):     (%.2f, %.2f)",
            params[[2]]$pars_fit[1],
            params[[2]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[2]]$L2_fit),
    
    sprintf("Beta model - unrestricted"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[3]]$pars_fit[1],
            params[[3]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[3]]$L2_fit),
    
    sprintf("Binormal model - concave"),
    sprintf("Parameters (alpha, beta):   (%.2f, %.2f)",
            params[[4]]$pars_fit[1],
            params[[4]]$pars_fit[2]
    ),
    sprintf("L2-distance:                 %.3f",
            params[[4]]$L2_fit)
  )
}) %>%
  cat(sep = "\n")

with(df_Vogel, {
  p <- plot_roc_empirical(empROC[[1]])
  p <- plot_binormal(params[[1]]$pars_fit, MDE_info[[1]], p, color = "red")
  p <- plot_binormal(params[[2]]$pars_fit, MDE_info[[2]], p, lty = 2, color = "red")
  p <- plot_beta(params[[3]]$pars_fit, MDE_info[[3]], p, color = "blue")
  p <- plot_beta(params[[4]]$pars_fit, MDE_info[[4]], p, lty = 2, color = "blue")
  p +
    theme(aspect.ratio = 1) +
    geom_segment(mapping = aes(x = 0, xend = 1, y = 0, yend = 1), lty = 2)
})

```

Empirical (black), fitted binormal (red) and fitted beta (blue) ROC curves in the unrestricted (solid) and concave (dashed) case.
